---
title: "Policy rate"
subtitle: "Banco de la República"
author: "Miguel Martinez"
date: today
format: 
  html:
    toc: true
    theme: cosmo
editor: visual
execute:
  warning: false
---

This R document presents the process for loading, munging and plotting the evolution of the daily daily-frequency of the policy rate of the Central Bank of Colombia, [Banco de la República](https://www.banrep.gov.co/es). The information is published by Banrep and you can download it [here](https://suameca.banrep.gov.co/buscador-de-series/#/).

The first step is to load the following libraries:

```{r}

# Load libraries

library(readxl)
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load data

Now we can load the data:

```{r}

# Load data

data = readxl::read_excel("graficador_series.xlsx",
                          sheet = "Datos",
                          range = "A1:I10133",
                          col_names = TRUE,
                          na = "-",
                          .name_repair = make_clean_names)
```

## Indexing

```{r}

# Let's omit the first row

data = data[-c(1), ]

```

## Transformation

Now we transform the date column with `as.Date`function:

```{r}

# Parsing to date

data = data |> 
  dplyr::mutate(
    fecha = as.Date(fecha, format = "%d/%m/%Y"))

```

Now, we can create the new columns: `year` `month` and `day`, from the `date` variable with the `lubridate` functions:

```{r}

# Extract information

data$año = lubridate::year(data$fecha)
data$mes = lubridate::month(data$fecha)
data$dia = lubridate::day(data$fecha)

glimpse(data)

```

```{r}

# Converting Tasa diaria to 'double' format replacing "," -> "."

data = data |>
  dplyr::select(fecha, año,
                mes, dia,
                tasa_de_politica_monetaria_dato_diario) |> 
  dplyr::mutate(
    tasa_de_politica_monetaria_dato_diario = 
      as.numeric(str_replace_all(
        string = tasa_de_politica_monetaria_dato_diario,
        pattern = ",",
        replacement = ".")))

data
```

## Filtering

Now we can filter the database for the first month of every year starting from 2015:

```{r}

# Filter the data

data = data |> 
  filter(año >= 2015)
```

## Graph

```{r}

# Create the plot

data |> 
ggplot(aes(x=fecha, y=tasa_de_politica_monetaria_dato_diario)) +
  geom_step(color="darkblue") +
  scale_x_date(date_breaks = "1 year", date_labels = "%m-%Y") +
  scale_y_continuous(breaks = seq(0,14, by = 2)) +
  theme_bw() +
  labs(title = "Tasa de intervención de política monetaria",
       subtitle = "(periodicidad diaria)",
       x = "Mes-Año",
       y = "Tasa (%)",
       caption = "Fuente: Elaboración propia con datos del Banco de la República, TIP") +
  theme(plot.caption.position = "panel",
        plot.caption = element_text(hjust = 0))
```

## Conclusion

In the previous graph we can see the latest monetary policy decisions by the Banco de la República regarding the REPO rate. Since the constitutional instruction is to maintain the purchasing power of the different agents of the economy and the inflation expectations anchored to an interval, we can observe that the REPO rate was reduced to get resources to finance the pandemic, and then quickly escalated in order to contain the raising inflation after the pandemic. It is worth mentioning that the [latest minutes](https://www.banrep.gov.co/es/noticias/minutas-banrep-octubre-2025) report discretion about reducing very quickly the interest rate levels, due to inflation levels, inflation expectations and external financing conditions.
